---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


##Loading and pre-processing the data: Getting the right packages
```{r, echo=TRUE}
library(sqldf)
library(plyr)
library(ggplot2)
library(dplyr)
```
###Setting up the right working directory and importing all the packages. Removed "NA" and converted date as date.
```{r, echo=TRUE}
setwd("D:/Coursera/Data Scientist/Reproducilble research/Week-1/Data")
activity<- read.csv("activity.csv")
activity1<-na.omit(activity)
activity1$date<-as.Date(activity1$date, "%Y-%m-%d")
```
## What is mean total number of steps taken per day? 
###Here used the sql approach. Grouped by date to calculate the sum. The histogram shows that there are more days where people walk anywhere between 10000 to 15000 steps.
```{r, echo=FALSE}
dailysteps_total<-sqldf('select sum(steps) as totaldlystps, date as date from activity1  group by date' )
hist(dailysteps_total$totaldlystps, main="Histogram of Daily steps", xlab = "Num of steps", ylab ="Freq(days)" )
```


## What is the average daily activity pattern? 

```{r, echo=TRUE}
StepsPer5min<-summarise(group_by(activity1,interval), steps=mean(steps))
g<-ggplot(StepsPer5min, aes(interval,steps))
print(g+ geom_line())
```
### The graph indicates that person takes maximum steps on an average at the 800+ interval
###following Code calculated the sum, mean and median for each day.
### It indicate that the average steps taken and the median
```{r, echo=FALSE}
agg<-summarize_each(group_by(activity1, date), funs(sum,mean, median), steps)
mean(activity1$steps)
median(activity1$steps)
```

## Imputing missing values
### for imputing the missing values I calculate the mean across each 5 min interval
```{r, echo=TRUE}
StepsPer5min<-summarise(group_by(activity1,interval), steps=mean(steps))
```
### then I join this data set based on the interval key back to the original 'unclean(no NA removed) data set
```{r, echo=TRUE}
activity2<-sqldf('select a.steps, a.date, a.interval, b.steps as avgstps from activity as a, StepsPer5min as b where a.interval = b.interval')
```
### This helps me with aligning the interval mean against the original dataset
### I then use a simple for and if loop to update the missing values with mean where ever necessary
```{r, echo=TRUE}
for(i in 1:length(activity2$steps))
  {
  if(is.na(activity2[i,1]))
    {
    activity2[i,1]<-activity2[i,4]
    }

  }
```
### After the values are imputed we do get more observation 61 days instead of 53.
```{r, echo= TRUE}
dailysteps_total1<-sqldf('select sum(steps) as totaldlystps, date as date from activity2  group by date' )
hist(dailysteps_total1$totaldlystps, main="Histogram of Daily steps", xlab = "Num of steps", ylab ="Freq(days)" )
```
### The above histogram(post imputing the data shows that we now have more number of days(35 compared to 25) where the subject has taken between 10000 and 15000 steps.
## Are there differences in activity patterns between weekdays and weekends?
### Using the imputed dataset, I add a new variable day and 
```{r, echo=TRUE}
activity2$day<-weekdays(as.Date(activity2$date,"%Y-%m-%d"))
activity2$Isweekend<-"weekday"
for(i in 1:length(activity2$steps))
  {
  if(is.na(activity2[i,1]))
    {
    activity2[i,1]<-activity2[i,4]
    }
  else
  {
   print("OK")   
   }

  }

```